### 2021-09-13 plos_latex_template.tex
comments on the 1st rule by Batool

Think about the reasons why you want to create a more diverse conference and try not to treat diversity as a checklist, avoid using diversity and inclusion work for image purposes, and do not act as a benefactor or savior but as a genuine collaborator.
%andrea: this is here because we want to avoid readers to read as recipes and treat this less seriously than how it should (and then blaming the recipe). i am not sure of the phrasing or position in the text."genuine collaborator" is in the spirit of co-liberation/co-conspirator
Actively be an ally: if you understand your own privileges (unearned advantages given by society to some people but not all) \cite{}, you can help to work on end oppression in your conference.
%Natalia: I'll look for a reference for privilege definition.
Think about the groups that are usually unsupported or discriminated against in your community or event, and about the meaning of "diverse" in your context. Imagine the result you would like to see if you succeeded in organizing a diverse conference according to your vision. Would this translate into an even gender distribution in your speakers? Would it be representation of all races -especially Black people-- among the head organizers, speakers, and attendees? Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions? This vision should guide and help hold the organizing team accountable along the way.

@@

@@

Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes far outweigh any inconveniences. 
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. On the other hand, a diverse team plays an important role at creating a welcoming space because seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes (i.e., representation) matters (see Rule \ref{rule_inclusion}). %Representation is an important aspect of diversity and inclusion: seeing people like you, from your country, that speak your language is one of the best way to feel you belong

@@

@@

If the conference is in person, the venue must be accessible for people with mobility limitations. 
Accommodations such as a quiet space for neurodivergent people should also be provided.
Presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them. 
Regardless of the conference format --online, in person, or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound. 
Platforms for conference registration and abstract submission, websites, and chat platforms --if used --should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection). 
Captioning for presentations should be available in more than one language if possible.
These aspects should be tested well in advance of going live.

@@

Make sure that every selection committee --the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees --are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has an implicit biases, to recognize them, and try to counteract them. An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles. The regional and local communities in your field are also good sources to tap into. 
For example, for useR! 2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference.

@@

\section{Set fair registration rates and provide financial support when needed}

@@

\section{Do not expect the organizing team to work full time and for free}

>>>>

Depending on the field, scientific or professional communities can have very different compositions. Thinking about which subgroups have been historically underrepresented in your community, and thinking about which subgroups hold most privileges can help you guide the organization of a more inclusive and diverse event. Doing this exercise while you ...
Think about the reasons why you want to create a more diverse conference can avoid falling into the mistake of treating diversity as a checklist, or using diversity and inclusion work for image purposes., and do not act as a benefactor or savior but as a genuine collaborator.
%andrea: this is here because we want to avoid readers to read as recipes and treat this less seriously than how it should (and then blaming the recipe). i am not sure of the phrasing or position in the text."genuine collaborator" is in the spirit of co-liberation/co-conspirator
Actively be an ally: if you understand your own privileges (unearned advantages given by society to some people but not all) \cite{}, you can help to work on end oppression in your conference. Privilege is largely invisible to those who have it. Once we understand our own privileges, we can begin addressing it on an individual and institutional basis.
%Natalia: I'll look for a reference for privilege definition.
%marcela: I modified the paragraph a bit, but it needs more work to connect Natalia's and Andrea's ideas. It's just a suggestion, feel free to re-write it again.
%Batool: I tried making some paraphrasing - It's just a simple suggestion, not sure if it deliver the message you mean it.  I also think that we may need to define "Privilege" at the 1st mention in the 1st paragraph in this rule. Something like "Privilege is a set of unearned benefits enjoyed by a group, beyond what’s available to others."

Think about the groups that are usually unsupported or discriminated against in your community or event, and about the meaning of "diverse" in your context. Imagine the result you would like to see if you succeeded in organizing a diverse conference according to your vision. Would this translate into an even gender distribution in your speakers? Would it be representation of all races -especially Black people-- among the head organizers, speakers, and attendees? Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions? This vision should guide and help hold the organizing team accountable along the way.

%batool: There are a few references which define privilege and might be good to add here including: Friedman, S., O’Brien, D., & McDonald, I. (2021). Deflecting Privilege: Class Identity and the Intergenerational Self. Sociology, 55(4), 716–733. https://doi.org/10.1177/0038038520982225
%batool: Another one is: Crevani, L. (2019). Privilege in place: How organisational practices contribute to meshing privilege in place. Scandinavian Journal of Management, 35(2). https://doi.org/10.1016/j.scaman.2018.09.002
%batool: please note that I didn't read both entire papers I mentioned above.

@@

Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of  attendants. If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, this sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community. Some examples are the newbies sessions for first timers, mixers lead by Rladies or MiR, or a leisure activity that reunites subgroups with the same interests: arts, exercise, sports, movies, etc.

@@

Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes far outweigh any inconveniences. 
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. On the other hand, a diverse team plays an important role at creating a welcoming space because seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes (i.e., representation) matters (see Rule \ref{rule_inclusion}). This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees. This gave a sense of variety and an appeal of “at least one networking session for everyone”.

%Representation is an important aspect of diversity and inclusion: seeing people like you, from your country, that speak your language is one of the best way to feel you belong

@@

Take advantage of the online setting for new networking activities that can be inclusive for people with different socialization styles, and accessibility needs. Social events in presencial conferences are often the most exclusive, and they are a very common place where you have code of conduct violations. Having them directly translated into their virtual versions does not address the need for creating opportunities for people to meet and bond. An online event is a great opportunity to create spaces for people who do not enjoy in person networking. Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with team work like trivia, offer some events that can be passively enjoyed like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session. Respect people’s limits, preferences, and remember that “the usual” does not necessarily work with your community, and that one single networking activity will never serve the whole community. It is worth trying new and varied activities that might work with subgroups of attendees.

@@

If the conference is in person, the venue must be accessible for people with mobility limitations. 
Accommodations such as a quiet space for neurodivergent people should also be provided.
Presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them. 
Regardless of the conference format --online, in person, or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound. Do not forget about art shows and movies when making sure that the materials presented are accessible for all. 
Platforms for conference registration and abstract submission, websites, and chat platforms --if used --should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection). 
Captioning for presentations should be available in more than one language if possible.
These aspects should be tested well in advance of going live.

@@

Make sure that every selection committee --the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees --are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has an implicit biases, to recognize them, and try to counteract them. An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles. The regional and local communities in your field are also good sources to tap into. 
For example, for useR! 2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference.

@@

\section{Set fair registration rates and provide financial support when needed}
\label{rule_costs}

@@

\section{Do not expect the organizing team to work full time and for free}
\label{rule_pay_team}
